---
title: "Homework"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will create a table of growth rate and plot all cities in Washington on a map, using the Google Visualization API. To get there, please install the **googleVis** package and then follow the steps below:

1. Re-create the `dt` object from the class by reading in the OFM data. It contains population for all cities in WA.

    ```{r eval = FALSE}
dt <- fread("ofm_april1_population_final.csv")
```

2. Extract the columns with city name (Jurisdiction) and the column with 2018 and 2019 population. Store the object in `mypop`. Rename the 2018 column to PopBase and the 2019 column to "Pop".

3. Convert the PopBase and Pop into numeric type.

4. Create a column Growth as (Pop - PopBase)/PopBase * 100 rounded to 2 decimals.

5. Add a column Rank that ranks cities by Growth in decreasing order. Hint: Use the `frank()` function. Ranking in decreasing order can be achieved by passing negative values of the original column.

6. Add a column State_share computed as the share of Pop within the state.

7. Add a column County_share computed as the share of Pop within its county. Hint: You need to compute sum(Pop) by counties first, assign it as an additional column to `mypo`, and then divide Pop by that.

8. Create a column "City" which takes values of "Jurisdiction" but removes any text containing " (part)". Hint: For replacing text use the function `gsub(" (part)", "", Jurisdiction, fixed = TRUE)` on the `j` dimension (just assign that call to `City`).

9. Create a google table by 

    ```{r eval = FALSE}
library(googleVis)
gtbl <- gvisTable(mypop[, .(Rank, County, City, Growth, PopBase, Pop, County_share, State_share)])
plot(gtbl)
```

10. The file "citycenters.csv" provided with the class data contains the latitude and longitude of cities in WA, which we will need for the map. Read the file into R using `fread` and store it in the object `cc`.

11. Rename the column "NAME" in the `cc` dataset to "City".

12. Join the datasets `mypop` and `cc` on the "City" column, so that columns "lat" and "long" appear as a new character column (called "latlong") of `mypop` in the form lat:long. Hint: Use the `paste` function to combine "lat" and "long" with `sep=":"`. In the joining command, don't forget to use the prefix `i.` for those two columns. 

13. Can you extract records where there was no match? Hint: the "latlong" column is NA for those records. Remove those records from `mypop`.

14. Create a column "Tip" constructed as `paste0(City, ": ", Pop, ", Gr: ", Growth)`.

15. Plot points on a map using:
    ```{r eval = FALSE}
plot(gvisMap(mypop, "latlong", "City", tipvar = "Tip", options = list(height = 800, width = 800)))
```

16. To combine a geo chart with a table on one page, you can use the following code:
    ```{r eval = FALSE}
gchart <- gvisGeoChart(mypop, "latlong", hovervar = "Tip", colorvar = "Pop", 
            options = list(region="US-WA", resolution = "metros", 
                           height = 800, width = 800))
plot(gvisMerge(gchart, gtbl))
```
